---
title: "Assignment 5"
output: html_notebook
---

# load packages
```{r}
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(rstatix)
library(ggpubr)
library(datarium)
```
# import data
```{r}
assignment5 = read_csv("assignment5_data.csv")
```
# Question 2
```{r}
subject_rt = assignment5 %>%
  group_by(subject, prime_condition) %>% 
  summarise(subject_rt = mean(response_RT))
    
subject_accuracy = assignment5 %>%
  group_by(subject, prime_condition) %>%
  summarise(subject_accuracy = mean(accuracy))
```

# Question 3
Both subject_accuracy and subject_rt are in long format because each subject has multiple rows.

# Question 4: wide
```{r}
subject_accuracy_wide = subject_accuracy %>%
  pivot_wider(names_from = prime_condition, values_from = subject_accuracy)
```

# Question 5: long
```{r}
subject_accuracy_long = subject_accuracy_wide %>%
  pivot_longer(names_to = "prime_condition", cols = both:unrelated)
```

# Question 6: Interpretation
subject_accuracy has the same information as subject_accuracy_long

# Question 7: t-test in R
```{r}
t.test(subject_accuracy_wide$phonological, subject_accuracy_wide$semantic, paired = TRUE)
```

# Question 8: t-test interpretation
answer

# Question 9: t-test manual
```{r}
x1_bar = mean(subject_accuracy_wide$phonological)
x2_bar = mean(subject_accuracy_wide$semantic)

s1 = sd(subject_accuracy_wide$phonological)
s2 = sd(subject_accuracy_wide$semantic)

n1 = nrow(subject_accuracy_wide)
n2 = subject_accuracy_wide %>% pull(semantic) %>% length()
```

```{r}
t_numerator = x1_bar - x2_bar
inside_root1 = s1*s1/n1
inside_root2 = s2*s2/n2
t_denominator = sqrt(inside_root1 + inside_root2)
t = t_numerator/t_denominator
```


```{r}
x = s1*s1/n1
y = s2*s2/n2

a = s1*s1*s1*s1/(n1*n1*(n1-1))
b = s2*s2*s2*s2/(n2*n2*(n2-1))

deg = (x + y)*(x+y)/(a + b)

p_value = 2*(1-pt(t,deg))
```

# Question 10: t-test outliers
```{r}
subject_accuracy_wide = subject_accuracy_wide %>%
  mutate(diff = phonological - semantic)

subject_accuracy_wide = subject_accuracy_wide %>% ungroup()
outliers = subject_accuracy_wide %>% identify_outliers(diff)
```
There is one outlier.

```{r}
hist(subject_accuracy_wide$diff)
```

```{r}
outlier_subs = outliers %>% pull(subject)

newdf = subject_accuracy_wide %>% filter(!subject %in% outlier_subs)
```

# Question 11
```{r}
ggqqplot(subject_accuracy_wide, "diff")
```

```{r}
subject_accuracy_wide %>% shapiro_test(diff)
```
The data does not significantly deviate from a normal distribution since the p-value was > 0.05. Normality is assumed.

# Question 12
The test of assumptions does not change my opinion on the validity of the t-test. Only one outlier was found but it was not extreme. Normality was assumed with the original data as well.
```{r}

```

